Supplementary Information, Possession trance covaries with measures of social rigidity in the Ethnographic Atlas

### 2023-02-01 Readme.Rmd
tidyup

---
title: Identifying covariate traits of trance and possession phenomena in the Ethnographic
  Atlas using boosting and nested regression
author: "Péter Rácz"
date: "`r format(Sys.Date(),'%e %B, %Y')`"
output:
  md_document:
    variant: markdown_github
bibliography: bibliography.bib
---

@@

This is the online supplement for the paper “Possession trance co-varies
with measures of social rigidity in the Ethnographic Atlas 2022”. The
current version of the paper is
[here](https://docs.google.com/document/d/1qbcsCeQ-Tw2W-yhQvvYd1Nk5ImQznqWBmiHX09lB_24/edit?usp=sharing). The `helper` file creates the data files. The `model` files run the Gradient Boosting Model and the Bayestraits model, respectively.

We have three questions. What sort of social factors have been proposed to co-occur with trance and possession phenomena? Can we give support to these using the Ethnographic Atlas and the Standard Cross-Cultural Sample? Using phylogenetic comparative methods, can we make claims on the causative structure between the co-occurrent factors?

Ee expect any effects we find to be relatively modest. This is because trance and possession phenomena result from complex interchanges between individual psychopathology and social organisation, and that, even if we can recognise broader patterns, the effects of social rigidity and marginalisation will be idiosyncratic on the ground. This means that both the outcome (trance and possession) and the predictors (e.g. measures of social rigidity) will be much more inherently variable than in similar analyses.

@@

```{r fig.width=18, fig.height=5}
lld = ld %>% 
  filter(var_id == 'EA112') %>% 
  mutate(
    lon2 = ifelse(lon < -20, lon + 360, lon),
    var_short = case_when(
      code == 1 ~ 'T',
      code == 2 ~ 'P',
      code == 3 ~ 'T,P,T iff P',
      code == 4 ~ 'T,P,T if P',
      code == 5 ~ 'T,P,T and P',
      code == 6 ~ 'P,T if P',
      code == 7 ~ 'T,P if T',
      code == 8 ~ 'none'
    ),
    coding = fct_reorder(var_name,code)
  )

p1 = ggplot() + 
  geom_polygon(data = mapWorld, aes(x=long, y = lat, group = group)) + 
  geom_label(data = lld, aes(x = lon2, y = lat, colour = coding, label = code)) +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks=element_blank(), axis.title.y=element_blank(), axis.text.y=element_blank()) +
  scale_colour_viridis_d() +
  theme_void()

p1 + plot_layout(guides = 'collect')
```

@@

Models are fit using Markov Chain Monte Carlo using exp 10 priors, 100 stepping stones, and 1000 iterations per stone. Each run provides a marginal likelihood for the model fit.

@@

```{r }
ld %>% 
  filter(var_id != 'EA112') %>% 
  distinct(var_id,var_title,var_description,recoded_value) %>% 
  arrange(var_id,recoded_value) %>% 
  kable('simple')
```

@@

```{r, echo=T}
gbm_params1 = list(
  learn_rate = c(0.01,0.02,0.03,0.1),
  max_depth = c(1,3,5,7),
  sample_rate = c(0.4,0.7,0.9,0.95),
  col_sample_rate = c(0.1,0.2,0.3),
  min_split_improvement = c(1e-3,1e-5)
)
```

>>>>

---
title: Supplementary Information, Possession trance covaries with measures of social rigidity in the Ethnographic Atlas 
author: "Péter Rácz"
date: "`r format(Sys.Date(),'%e %B, %Y')`"
output:
  md_document:
    variant: markdown_github
bibliography: bibliography.bib
---

@@

This is the online supplement for the paper “Possession trance covaries
with measures of social rigidity in the Ethnographic Atlas”. The
current version of the paper is
[here](https://docs.google.com/document/d/1qbcsCeQ-Tw2W-yhQvvYd1Nk5ImQznqWBmiHX09lB_24/edit?usp=sharing). The `helper` file creates the data files. The `model` files run the Gradient Boosting Model and the Bayestraits model, respectively.

We have three questions. What sort of social factors have been proposed to cooccur with trance and possession phenomena? Can we give support to these using the Ethnographic Atlas and the Standard Cross-Cultural Sample? Using phylogenetic comparative methods, can we make claims on the causative structure between the cooccurrent factors?

We expect any effects we find to be relatively modest. This is because trance and possession phenomena result from complex interchanges between individual psychopathology and social organisation, and that, even if we can recognise broader patterns, the effects of social rigidity and marginalisation will be idiosyncratic on the ground. This means that both the outcome (trance and possession) and the predictors (e.g. measures of social rigidity) will be much more inherently variable than in similar analyses.

@@

```{r fig.width=8, fig.height=9}
lld = ld %>% 
  filter(var_id == 'EA112') %>% 
  mutate(
    lon2 = ifelse(lon < -20, lon + 360, lon),
    var_short = case_when(
      code == 1 ~ 'trance',
      code == 2 ~ 'possession',
      code == 3 ~ 'trance, possession,\ntrance iff possession',
      code == 4 ~ 'trance, possession,\ntrance if possession',
      code == 5 ~ 'trance, possession,\ntrance and possession',
      code == 6 ~ 'possession,\ntrance if possession',
      code == 7 ~ 'trance,\npossession if trance',
      code == 8 ~ 'none'
    ),
    coding = fct_reorder(var_name,code),
    var_short = fct_reorder(var_short,code)
  )

p1 = ggplot() + 
  geom_polygon(data = mapWorld, aes(x=long, y = lat, group = group)) + 
  # geom_label(data = lld, aes(x = lon2, y = lat, label = ' ')) +
  geom_point(data = lld, aes(x = lon2, y = lat), colour = '#FFAA00') +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks=element_blank(), axis.title.y=element_blank(), axis.text.y=element_blank()) +
  # scale_colour_viridis_d() +
  # guides(colour = 'none',label = 'none') +
  theme_void() +
  facet_wrap( ~ var_short, nrow = 4)

p1
```

@@

Models are fit using Markov Chain Monte Carlo using exp(10) priors, 100 stepping stones, and 1000 iterations per stone. Each run provides a marginal likelihood for the model fit.

@@

```{r }
coding_table = ld %>% 
  filter(var_id != 'EA112') %>% 
  distinct(var_id,var_title,var_description,recoded_value) %>% 
  arrange(var_id,recoded_value)

coding_table %>% 
  kable('simple')

# write_sheet(coding_table, ss = 'https://docs.google.com/spreadsheets/d/1X7ZCnw6LVhjLIy4SiSeK3iZ6o1S144iHMv4DqOaCDEU/edit#gid=0', sheet = 'Table 9')

@@

```{r, echo=T}
gbm_params1 = list(
  learn_rate = c(0.01,0.02,0.03,0.1),
  max_depth = c(1,3,5,7),
  sample_rate = c(0.4,0.7,0.9,0.95),
  col_sample_rate = c(0.1,0.2,0.3),
  min_split_improvement = c(1e-3,1e-5)
)
